---
title: "Classification with linear Regression and K-NN"
output:
  flexdashboard::flex_dashboard:
    navbar:
    - align: right
      href: https://web.stanford.edu/~hastie/ElemStatLearn/
      title: More Help
    orientation: columns
    vertical_layout: fill
resource_files:
- data/zip_test.gz
- data/zip_train.gz
- data/zip_test.gz
- data/zip_train.gz
- data/zip_train.gz
- data/zip_test.gz
- data/zip_train.gz
- data/zip_test.gz
runtime: shiny
autor: Akbar Akbari Esfahani
---



Linear Regression Classification
=======================================================================

```{r setup, include=FALSE}
#install.packages("nlme")
library(flexdashboard)
library(tidyverse)
library(ggfortify)
library(class)
library(caret)
library(PerformanceAnalytics)
library(shiny)
require(e1071)
```

Input {.sidebar}
-----------------------------------------------------------------------

### Input Selection
```{r}
numericInput('num_1', 'Please select first digit',
             value = 0, min = 0, max = 10, step = 1)
numericInput('num_2', 'Please select second digit',
             value = 0, min = 0, max = 10, step = 1)
actionButton("go", "Go")   
```



Column {data-width=650}
-----------------------------------------------------------------------

### Decision Boundary / Hyper-plane seperation

```{r}
# function to exract any two digits for classification
digit_subset <- function(dig_1, dig_2, data_in) {
          data_in %>%
               mutate(class = ifelse(V1 == dig_1, dig_1, 
                                     ifelse(V1 == dig_2, dig_2,
                                            "other"))) %>%
               dplyr::filter(class != "other") %>%
               select(-V1)
     }
     
randomVal <- eventReactive(input$go, {
     data.frame(num_1 = input$num_1, num_2 = input$num_2)
})

train_d <- reactive({
     randomVal1 <- randomVal()
     num_1 <- randomVal1$num_1
     num_2 <- randomVal1$num_2
     train_data <- digit_subset(num_1, num_2, 
     read.table("zip_train.gz"))
})
test_d <- reactive({
     randomVal1 <- randomVal()
     num_1 <- randomVal1$num_1
     num_2 <- randomVal1$num_2
     test_data <- digit_subset(num_1, num_2, 
     read.table("zip_test.gz"))
})



renderPlot({
        train_data <- train_d()
        ans <- train_data %>%
             select(-class) %>%
             princomp(.)
        
        # lets look at the distribution of 7 versus other numbers in the 
        #  first two principle components
        autoplot(prcomp(train_data[,-257]), data = train_data, colour = "class") +
             xlab("First Principle Component") +
             ylab("Second Principle Component") 
   })
plotOutput("plot")

# lets look at the distribution of 7 in the first three principle components     

```

Column {data-width=350}
-----------------------------------------------------------------------

### Summary of Linear Regression Fit

```{r}
# linear Regression model for binary class ------------------------------------------------------------------------
renderPrint({
     train_data <- train_d()
     lm_fit <- lm(class ~ ., train_data)
     summary(lm_fit)
})



# prediction based on linear model --------------------------------------------------------------------------------



```

### Summary Statistics on Predicion of Test Set

```{r}
renderPrint({
     randomVal1 <- randomVal()
     num_1 <- randomVal1$num_1
     num_2 <- randomVal1$num_2
     
     train_data <- train_d()
     test_data <- test_d()
     lm_fit <- lm(class ~ ., train_data)
     avrg <- (num_1 + num_2)/2
     testing <- test_data %>%
          select(-class)
     
     test_results <- lm_fit %>% 
          predict.lm(., newdata=testing)

     test_results <- data.frame(test_results) %>%
     mutate(returned = 1*(test_results > avrg))
     
     te.er <- test_data %>%
          mutate(bin_classes = 1*(class > as.character(avrg))) %>%
          select(bin_classes) %>%
          .$bin_classes
     temp_er_table <- table(te.er, test_results$returned)
     caret::confusionMatrix(temp_er_table)


     # error.test = (temp_er_table[1,2] + temp_er_table[2,1])/
     # length(test_results$returned)
     # error.test
})
#temp_er_table
```

K-Nearest Neighbor Classification
=======================================================================
Column {data-width=650}
-----------------------------------------------------------------------

### Principle Component Plane Class Distribution

```{r}
# function to exract any two digits for classification
renderPlot({
     train_data <- train_d()
     test_data <- test_d()
     train.dat <- train_data %>%
             select(-class)
     test.dat  <- test_data %>%
          select(-class)
     Y <- train_data$class
     Y.test <- test_data$class
     ###########  k = 1 ###########
ans.knn.1 <- knn(train = train.dat, test=train.dat, 
                 k=1, cl=Y)
table(Y, ans.knn.1)
ans.knn.1 <- knn(train = train.dat, test=test.dat, 
                 k=1, cl=Y)
k.t.e1 = table(Y.test, ans.knn.1)
er.t.knn1 = (k.t.e1[1,2]+k.t.e1[2,1])/length(Y.test)


###########  k = 3 ###########
ans.knn.3 <- knn(train = train.dat, test=train.dat, 
                 k=3, cl=Y)
k.t.t3 = table(Y, ans.knn.3)
er.tr.knn3 = (k.t.t3[1,2]+k.t.t3[2,1])/length(Y)
er.tr.knn3

ans.knn.3 <- knn(train = train.dat, test=test.dat, 
                 k=3, cl=Y)
k.t.e3 = table(Y.test, ans.knn.3)
er.t.knn3 = (k.t.e3[1,2]+k.t.e3[2,1])/length(Y.test)
er.t.knn3

###########  k = 5 ###########
ans.knn.5 <- knn(train = train.dat, test=train.dat, 
                 k=5, cl=Y)
k.t.t5 = table(Y, ans.knn.5)
er.tr.knn5 = (k.t.t5[1,2]+k.t.t5[2,1])/length(Y)
er.tr.knn5


ans.knn.5 <- knn(train = train.dat, test=test.dat, 
                 k=5, cl=Y)
k.t.e5 = table(Y.test, ans.knn.5)
er.t.knn5 = (k.t.e5[1,2]+k.t.e5[2,1])/length(Y.test)
er.t.knn5


###########  k = 7 ###########
ans.knn.7 <- knn(train = train.dat, test=train.dat, 
                 k=7, cl=Y)
k.t.t7 = table(Y, ans.knn.7)
er.tr.knn7 = (k.t.t7[1,2]+k.t.t7[2,1])/length(Y)
er.tr.knn7

ans.knn.7 <- knn(train = train.dat, test=test.dat, 
                 k=7, cl=Y)
k.t.e7 = table(Y.test, ans.knn.7)
er.t.knn7 = (k.t.e7[1,2]+k.t.e7[2,1])/length(Y.test)
er.t.knn7

###########  k = 15 ###########
ans.knn.15 <- knn(train = train.dat, test=train.dat, 
                  k=15, cl=Y)
k.t.t15 = table(Y, ans.knn.15)
er.tr.knn15 = (k.t.t15[1,2]+k.t.t15[2,1])/length(Y)
er.tr.knn15
 
ans.knn.15 <- knn(train = train.dat, test=test.dat, 
                  k=15, cl=Y)
k.t.e15 = table(Y.test, ans.knn.15)
er.t.knn15 = (k.t.e15[1,2]+k.t.e15[2,1])/length(Y.test)
er.t.knn15


# let's put this together in a plot like Figure 2.4 of the text

train <- c(0, 9, 11, 19, 30)/1289
test  <- c(8, 8, 8, 9, 12)/324
k <- c(1,3,5,7,15)
reg.test  <- 14/324
reg.train <- 12/1289

plot(k, test, xlab="k", ylab="", type="b",
     lty=1, col=1, pch="x", ylim=c(0,0.05),
     main="Error Rate as a Function of k")
lines(k, train, lty=2, col=2, pch="o", type="b")
lines(c(1,15), c(reg.test, reg.test), lty=3, col=3)
lines(c(1,15), c(reg.train, reg.train), lty=4, col=4)
legend(9, .03, c("NN - Test Data","NN - Training Data", "Reg - Test", "Reg - Train"),
       lty=c(0,0,3,4), col=1:4, pch=c("x","o","", ""))

})

# lets look at the distribution of 7 in the first three principle components     

```

Column {data-width=350}
-----------------------------------------------------------------------

### Summary Statistics on Predicion of Test Set with K-NN

```{r}
# linear Regression model for binary class ------------------------------------------------------------------------
renderPrint({
     train_data <- train_d()
     test_data <- test_d()
     train.dat <- train_data %>%
             select(-class)
     test.dat  <- test_data %>%
          select(-class)
     Y <- train_data$class
     Y.test <- test_data$class
     ###########  k = 1 ###########
ans.knn.1 <- knn(train = train.dat, test=train.dat, 
                 k=1, cl=Y)
table(Y, ans.knn.1)
ans.knn.1 <- knn(train = train.dat, test=test.dat, 
                 k=1, cl=Y)
k.t.e1 = table(Y.test, ans.knn.1)

     caret::confusionMatrix(k.t.e1 )
})



# prediction based on linear model --------------------------------------------------------------------------------



```

### Summary Statistics on Predicion of Test Set with Regression

```{r}
renderPrint({
     randomVal1 <- randomVal()
     num_1 <- randomVal1$num_1
     num_2 <- randomVal1$num_2
     
     train_data <- train_d()
     test_data <- test_d()
     lm_fit <- lm(class ~ ., train_data)
     avrg <- (num_1 + num_2)/2
     testing <- test_data %>%
          select(-class)
     
     test_results <- lm_fit %>% 
          predict.lm(., newdata=testing)

     test_results <- data.frame(test_results) %>%
     mutate(returned = 1*(test_results > avrg))
     
     te.er <- test_data %>%
          mutate(bin_classes = 1*(class > as.character(avrg))) %>%
          select(bin_classes) %>%
          .$bin_classes
     temp_er_table <- table(te.er, test_results$returned)
     caret::confusionMatrix(temp_er_table)


     # error.test = (temp_er_table[1,2] + temp_er_table[2,1])/
     # length(test_results$returned)
     # error.test
})
#temp_er_table
```

Data Help and Notes
=======================================================================

In this app we will compare the classification performance of linear regression and k–nearest neighbor classification on handwritten zipcode data. Data can be found here:  www-stat.stanford.edu/ElemStatLearn

Data Explantion: 
Normalized handwritten digits, automatically scanned from envelopes by the U.S. Postal Service. The original scanned digits are binary and of different sizes and orientations; the
images  here have been deslanted and size normalized, resulting in 16 x 16 grayscale images (Le Cun et al., 1990).

The data are in two gzipped files, and each line consists of the digit id (0-9) followed by the 256 grayscale values.
  
Alternatively, the training data are available as separate files per digit (and hence without the digit identifier in each row)  

The test set is notoriously "difficult", and a 2.5% error rate is excellent. These data were kindly made available by the neural network group at AT&T research labs (thanks to Yann Le Cunn). 

There are 7291 training observations and 2007 test observations,
distributed as follows:

```{r}
renderTable({
     data.frame("Number_0" = c(1194, 359), "Number_1" = c(1005, 264),
                "Number_2" = c(731, 198), "Number_3" = c(658, 166), 
                "Number_4" = c(652, 200), "Number_5" = c(556, 160),
                "Number_6" = c(664, 170), "Number_7" = c(645, 147),
                "Number_8" = c(542, 166), "Number_9" = c(644, 177),
                "Total" = c(7291, 2007), row.names = c("Train", "Test"))
})

```
 